---
title: "P1_Trachea_10X_Dec11_mm10.1.2.0"
output: html_notebook
---
## P1 trachea sample on Dec11 2018. v2 chemistry
```{r}
library(Seurat)
library(dplyr)
```
```{r}
setwd("./Trachea_Dec_mm10.1.2.0")
# Load data
file_10Xdir_P1_Dec11<-c("P1_Dec11_mut_1","P1_Dec11_mut_2","P1_Dec11_wt_6","P1_Dec11_wt_7")
names(file_10Xdir_P1_Dec11)<-c("P1_Dec11_mut_1","P1_Dec11_mut_2","P1_Dec11_wt_6","P1_Dec11_wt_7")
P1_Dec11.data <- Read10X(data.dir = file_10Xdir_P1_Dec11)
```
```{r}
P1_Dec11_mm10.1.2.0_Trachea <- CreateSeuratObject(raw.data = P1_Dec11.data, min.cells = 1, min.genes = 1000, 
    project = "allP1_10X_Trachea")
```
```{r}
P1_Dec11_mm10.1.2.0_Trachea <- FilterCells(object = P1_Dec11_mm10.1.2.0_Trachea, subset.names = c("nGene", "nUMI"), 
    low.thresholds = c(1000, 4000), high.thresholds = c(Inf, Inf))
P1_Dec11_mm10.1.2.0_Trachea@data@Dim
```
```{r}
ribo.genes <- grep(pattern = "^Rp[sl][[:digit:]]", x = rownames(x = P1_Dec11_mm10.1.2.0_Trachea@data), value = TRUE)
percent.ribo <- Matrix::colSums(P1_Dec11_mm10.1.2.0_Trachea@raw.data[ribo.genes, ])/Matrix::colSums(P1_Dec11_mm10.1.2.0_Trachea@raw.data)
P1_Dec11_mm10.1.2.0_Trachea <- AddMetaData(object = P1_Dec11_mm10.1.2.0_Trachea, metadata = percent.ribo, col.name = "percent.ribo")
```

```{r}
mito.genes <- grep(pattern = "^mt-", x = rownames(x = P1_Dec11_mm10.1.2.0_Trachea@data), value = TRUE)
percent.mito <- Matrix::colSums(P1_Dec11_mm10.1.2.0_Trachea@raw.data[mito.genes, ])/Matrix::colSums(P1_Dec11_mm10.1.2.0_Trachea@raw.data)
P1_Dec11_mm10.1.2.0_Trachea <- AddMetaData(object = P1_Dec11_mm10.1.2.0_Trachea, metadata = percent.mito, col.name = "percent.mito")
```
```{r}
cell_name<-read.table(text=P1_Dec11_mm10.1.2.0_Trachea@cell.names,sep="_",colClasses = "character")
age<-cell_name[,1]
names(age)<-P1_Dec11_mm10.1.2.0_Trachea@cell.names
```
```{r}
P1_Dec11_mm10.1.2.0_Trachea<-AddMetaData(object = P1_Dec11_mm10.1.2.0_Trachea, metadata = age, col.name = "age")
```

```{r}
sample_date<-cell_name[,2]
names(sample_date)<-P1_Dec11_mm10.1.2.0_Trachea@cell.names
```
```{r}
P1_Dec11_mm10.1.2.0_Trachea<-AddMetaData(object = P1_Dec11_mm10.1.2.0_Trachea, metadata = sample_date, col.name = "sample_date")
```

```{r}
genotype<-cell_name[,3]
names(genotype)<-P1_Dec11_mm10.1.2.0_Trachea@cell.names
```
```{r}
P1_Dec11_mm10.1.2.0_Trachea<-AddMetaData(object = P1_Dec11_mm10.1.2.0_Trachea, metadata = genotype, col.name = "genotype")
```


```{r}
seq_group<-paste(cell_name[,1],cell_name[,2],cell_name[,3],cell_name[,4],sep="_")
names(seq_group)<-P1_Dec11_mm10.1.2.0_Trachea@cell.names
```
```{r}
P1_Dec11_mm10.1.2.0_Trachea<-AddMetaData(object = P1_Dec11_mm10.1.2.0_Trachea, metadata = seq_group, col.name = "seq_group")
```


```{r}
table(P1_Dec11_mm10.1.2.0_Trachea@meta.data$seq_group)
```
##### add doublet score for each cell as metadata:
```{r}
load("./P1_Dec11/mm10-1.2.0/P1_Dec11_doubletScore.RData")
P1_Dec11_mm10.1.2.0_Trachea<-AddMetaData(object = P1_Dec11_mm10.1.2.0_Trachea, metadata = P1_Dec11_mm10.1.2.0_Trachea, col.name = "doublet_score")
sum(is.na(P1_Dec11_mm10.1.2.0_Trachea@meta.data$doublet_score))
```


```{r}
aggregate(P1_Dec11_mm10.1.2.0_Trachea@meta.data[, c(1:2,9:10)], list(P1_Dec11_mm10.1.2.0_Trachea@meta.data$seq_group), median)

```

```{r}
P1_Dec11_mm10.1.2.0_Trachea <- ScaleData(object = P1_Dec11_mm10.1.2.0_Trachea)
```


```{r}
P1_Dec11_mm10.1.2.0_Trachea <- FindVariableGenes(object = P1_Dec11_mm10.1.2.0_Trachea, do.plot = TRUE, x.low.cutoff=0.1,x.high.cutoff = Inf, y.cutoff = 0.5)
```
```{r}
P1_Dec11_mm10.1.2.0_Trachea <- RunPCA(object = P1_Dec11_mm10.1.2.0_Trachea,pcs.compute = 40, do.print = FALSE)
P1_Dec11_mm10.1.2.0_Trachea <- ProjectPCA(object = P1_Dec11_mm10.1.2.0_Trachea, do.print = FALSE)
```
```{r,fig.height=12,fig.width=14}
PCHeatmap(object = P1_Dec11_mm10.1.2.0_Trachea, pc.use = c(1:3,18:20,38:40), cells.use = 500, do.balanced = TRUE, label.columns = FALSE, num.genes = 30)

```
```{r}
PCElbowPlot(object = P1_Dec11_mm10.1.2.0_Trachea)
```
```{r}
n.pcs = 17
res.used <- 0.8

P1_Dec11_mm10.1.2.0_Trachea <- FindClusters(object = P1_Dec11_mm10.1.2.0_Trachea, reduction.type = "pca", dims.use = 1:n.pcs, 
    resolution = res.used, print.output = 0, save.SNN = TRUE)
```
```{r}
P1_Dec11_mm10.1.2.0_Trachea <- RunTSNE(object = P1_Dec11_mm10.1.2.0_Trachea, dims.use = 1:n.pcs, seed.use = 10, perplexity=30, dim.embed = 2)

```

```{r}
TSNEPlot(object = P1_Dec11_mm10.1.2.0_Trachea, do.label = T,group.by="res.0.8",pt.size = 0.2)

```


```{r}
res.used <- 1.2

P1_Dec11_mm10.1.2.0_Trachea <- FindClusters(object = P1_Dec11_mm10.1.2.0_Trachea, reduction.type = "pca", dims.use = 1:n.pcs, 
    resolution = res.used, print.output = 0, save.SNN = TRUE)
```
```{r}
P1_Dec11_mm10.1.2.0_Trachea <- RunTSNE(object = P1_Dec11_mm10.1.2.0_Trachea, dims.use = 1:n.pcs, seed.use = 10, perplexity=30, dim.embed = 2,k.param=15)

```

```{r}
TSNEPlot(object = P1_Dec11_mm10.1.2.0_Trachea, do.label = T,pt.size = 0.2,group.by="res.1.2")

```
```{r}
TSNEPlot(object = P1_Dec11_mm10.1.2.0_Trachea, do.label = F,group.by="seq_group",pt.size = 0.2)

```

```{r}
save(P1_Dec11_mm10.1.2.0_Trachea,file="seu_P1_Dec11_mm10.1.2.0.RData")
```

```{r}
load(file="seu_P1_Dec11_mm10.1.2.0.RData")
```
```{r,fig.height=12,fig.width=35}
DoHeatmap(object = P1_Dec11_mm10.1.2.0_Trachea, genes.use = c("Epcam","Trp63","Krt5","Mki67","Top2a","Creb3l1","Muc5ac","Gp2","Galnt6","Ptgdr","Cd177","Foxj1","Foxn4","Calca","Snap25","Tubb3","Ascl1","Chga","Cd3g","C1qb","Fcer1g","Mbp","Plp1","Egfl7","Cdh5","Pecam1","Nrp1","Lyve1","Cd34","Pi16","Akap12","Dcn","Galnt15","Cldn11","Col11a1","Lum","Acan","Wif1","Col2a1","Ano1","Acta2","Tagln","Myh11","Ppp1r14a","Notch3","Rgs5","Mia","Ednra","Sfrp2","Pdgfrb","Hhip","Sox9","Twist2"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,group.by = "res.1.2",group.order = c(9,0,7,4,8,15,14,13,3,10,11,12,5,1,2,6,16),cex.row = 20,group.cex = 30
  )
```


```{r}
df_P1_Dec11<-FetchData(P1_Dec11_mm10.1.2.0_Trachea,c("Areg","Il10","Il13","Il18","Il1b","Il33","Il4","Il5","Il6","S100a8","S100a9","Tgfb2","Tnf","Tslp","res.1.2","res.0.8","age","genotype","seq_group","Pigr","Cxcl17","Cxcl14","Ccl25","Ccl28","Cxcl15","Ccl20","Retnlb","Retnla","Itln1","Foxj1","Spdef","Creb3l1","Gp2","Tff2","Scgb1a1","Lcn2","Lgals3","Cftr","Foxi1"))

```


```{r}
table(P1_Dec11_mm10.1.2.0_Trachea@meta.data$res.0.8,P1_Dec11_mm10.1.2.0_Trachea@meta.data$res.1.2)
```
##### use resolution=1.2 for the rest of the analysis

```{r}
P1_Dec11_mm10.1.2.0_res12_5<-FindMarkers(P1_Dec11_mm10.1.2.0_Trachea,ident.1=c(5),only.pos = TRUE)
P1_Dec11_mm10.1.2.0_res12_5
```
```{r}
P1_Dec11_mm10.1.2.0_res12_12<-FindMarkers(P1_Dec11_mm10.1.2.0_Trachea,ident.1=c(12),only.pos = TRUE)
P1_Dec11_mm10.1.2.0_res12_12
```
```{r}
P1_Dec11_mm10.1.2.0_res12_13<-FindMarkers(P1_Dec11_mm10.1.2.0_Trachea,ident.1=c(13),only.pos = TRUE)
P1_Dec11_mm10.1.2.0_res12_13
```
```{r}
P1_Dec11_mm10.1.2.0_res12_3<-FindMarkers(P1_Dec11_mm10.1.2.0_Trachea,ident.1=c(3),only.pos = TRUE)
P1_Dec11_mm10.1.2.0_res12_3
```
```{r}
P1_Dec11_mm10.1.2.0_res12_3over2<-FindMarkers(P1_Dec11_mm10.1.2.0_Trachea,ident.1=c(3),ident.2 = c(2),only.pos = TRUE)
P1_Dec11_mm10.1.2.0_res12_3over2
```
```{r}
P1_Dec11_mm10.1.2.0_res12_1over3<-FindMarkers(P1_Dec11_mm10.1.2.0_Trachea,ident.1=c(1),ident.2 = c(3),only.pos = TRUE)
P1_Dec11_mm10.1.2.0_res12_1over3
```

##### basal, secretory, and ciliated:
```{r}
P1_Dec11_mm10.1.2.0_Trachea <- SetAllIdent(object = P1_Dec11_mm10.1.2.0_Trachea, id = "res.1.2")
P1_Dec11_epi<-SubsetData(object=P1_Dec11_mm10.1.2.0_Trachea,ident.use=c(0,7,4,8,9))
```
```{r}
colnames(P1_Dec11_epi@meta.data)[colnames(P1_Dec11_epi@meta.data) == 'res.0.8'] <- 'orig.0.8'
colnames(P1_Dec11_epi@meta.data)[colnames(P1_Dec11_epi@meta.data) == 'res.1.2'] <- 'orig.1.2'

P1_Dec11_epi <- ScaleData(object = P1_Dec11_epi)
```
```{r}
P1_Dec11_epi <- FindVariableGenes(object = P1_Dec11_epi, do.plot = TRUE, x.low.cutoff=0.1,x.high.cutoff = Inf, y.cutoff = 0.5)
```
######run PCA on the set of genes
```{r}
P1_Dec11_epi <- RunPCA(object = P1_Dec11_epi, do.print = FALSE)
PCAPlot(P1_Dec11_epi)
```

```{r}
P1_Dec11_epi <- ProjectPCA(object = P1_Dec11_epi, do.print = F)
```

```{r}
PCElbowPlot(object = P1_Dec11_epi)
```
```{r,fig.height=50,fig.width=15}
PCHeatmap(object = P1_Dec11_epi, pc.use = 1:20, cells.use = 500, do.balanced = TRUE, label.columns = FALSE, num.genes = 25)

```


```{r}
n.pcs.sub = 12
```
#####resolution parameter sets the ‘granularity’ of the downstream clustering, with increased values leading to a greater number of clusters. 
```{r}
res.used <- 1.2
```

```{r}
P1_Dec11_epi <- FindClusters(object = P1_Dec11_epi, reduction.type = "pca", dims.use = 1:n.pcs.sub, 
                     resolution = res.used, print.output = 0, force.recalc = T)
```
```{r}
P1_Dec11_epi <- RunTSNE(object = P1_Dec11_epi, dims.use = 1:n.pcs.sub, perplexity=30)
```
```{r, fig.width=10,fig.height=9}
TSNEPlot(object = P1_Dec11_epi, do.label = T,pt.size = 0.4)
```
```{r}
res.used <- 0.8
```

```{r}
P1_Dec11_epi <- FindClusters(object = P1_Dec11_epi, reduction.type = "pca", dims.use = 1:n.pcs.sub, 
                     resolution = res.used, print.output = 0, force.recalc = T)
```
```{r}
P1_Dec11_epi <- RunTSNE(object = P1_Dec11_epi, dims.use = 1:n.pcs.sub, perplexity=30)
```
```{r, fig.width=10,fig.height=9}
TSNEPlot(object = P1_Dec11_epi, do.label = T,pt.size = 0.4,group.by="res.0.8")
```
##### use resolution=0.8 for the rest of the analysis

```{r,fig.height=9,fig.width=35}

DoHeatmap(object = P1_Dec11_epi, genes.use = c("Epcam","Trp63","Krt5","Krt14","Mki67","Top2a","Krt4","Creb3l1","Muc5ac","Gp2","Galnt6","Ptgdr","Cd177","Foxj1","Foxn4","Spag17","Ccno","Shisa8"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,group.by = "res.0.8",cex.row = 25,group.cex = 30,group.order = c(5,7,3,6,0,4,1,8,2)
  )
```
##### c1 seems to be hybrid of ciliated and secretory cells.
```{r,fig.height=6,fig.width=24}
P1_Dec11_epi <- SetAllIdent(object = P1_Dec11_epi, id = "res.0.8")

DoHeatmap(object = P1_Dec11_epi, genes.use = c("Foxj1","Top2a","Mcidas","Ccno","Foxn4","Shisa8","Lrrc23","Prr18","Cfap53","Cdhr3","Sntn","Ifitm1","Lbp","Ly6c1","Ly6a","3300002A11Rik","Slc6a6","Xbp1","Creb3l1","Gp2","Muc5ac","Muc5b","Spdef"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use = P1_Dec11_epi@cell.names[P1_Dec11_epi@meta.data$res.1.2 %in% c(1,2,8)],group.order = c(8,2,1),group.cex = 30,cex.row = 20
  )
```


##### secretory cells:
```{r,fig.height=8,fig.width=34}
DoHeatmap(object = P1_Dec11_epi, genes.use = c("Epcam","Mki67","Top2a","Krt4","Krt13","Lgals3","Tppp3","Clic3","Sec14l3","Krt8","Krt18","Muc16","Creb3l1","Spdef","Gp2","Tff2","Muc5ac","Muc5b","Galnt6","Ptgdr","Cd177"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use = P1_Dec11_epi@cell.names[P1_Dec11_epi@meta.data$res.0.8 %in% c(3,0,6,4)],cex.row = 20,group.cex = 30,group.order = c(3,0,6,4)
  )
```
```{r}
marker_P1epi_res0.8_6over0<-FindMarkers(P1_Dec11_epi,ident.1=c(6),ident.2 = c(0),only.pos = TRUE)
marker_P1epi_res0.8_6over0
```
```{r}
marker_P1epi_res0.8_6over4<-FindMarkers(P1_Dec11_epi,ident.1=c(6),ident.2 = c(4),only.pos = TRUE)
marker_P1epi_res0.8_6over4
```



```{r,fig.width=5,fig.height=5}
ggplot(data=P1_Dec11_epi@meta.data[P1_Dec11_epi@meta.data$res.0.8 %in% c(3,0,4,6),],aes(genotype,fill=res.0.8))+ 
    geom_bar(position="fill")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r,fig.height=4,fig.width=15}
DoHeatmap(object = P1_Dec11_epi, genes.use = c("Epcam","Krt4","Lgals3","Tppp3","Creb3l1","Spdef","Gp2"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use = P1_Dec11_epi@cell.names[P1_Dec11_epi@meta.data$res.0.8 %in% c(4)],cex.row = 20,group.cex = 30,group.by = "genotype"
  )
```


```{r}
P1_Dec11_epi@meta.data$cell_type<-mapvalues(P1_Dec11_epi@meta.data$res.0.8,from=c("0","1","2","3","4","5","6","7","8"),to=c("Secretory","CiliaSecretory","Ciliated","Secretory","Secretory","Basal","Secretory","Basal/Secretory","Ciliated"))
```

```{r,fig.width=5,fig.height=5}
ggplot(data=P1_Dec11_epi@meta.data,aes(genotype,fill=cell_type))+ 
    geom_bar(position="fill")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r,fig.width=5,fig.height=5}
table(P1_Dec11_epi@meta.data$cell_type,P1_Dec11_epi@meta.data$genotype)
```


```{r}
DE_P1_secretory_genotype<-FindMarkers(P1_Dec11_epi,cells.1<-WhichCells(object=P1_Dec11_epi,cells.use = (P1_Dec11_epi@meta.data$genotype=="wt" & P1_Dec11_epi@meta.data$cell_type=="Secretory" )),cells.2<-WhichCells(object=P1_Dec11_epi,cells.use = (P1_Dec11_epi@meta.data$genotype=="mut" & P1_Dec11_epi@meta.data$cell_type=="Secretory" )),only.pos = F,logfc.threshold=0,min.pct=0.05)
DE_P1_secretory_genotype
```
```{r}
write.table(DE_P1_secretory_genotype,"DE_P1_secretory_genotype.txt",sep="\t")
```

```{r}
#P1_sec_geneList<-c("Krt4","Lyz1","Lcn2","Itln1","Cxcl15","Retnla","Tsc22d3","Cxcl17","Atp1b1","Tppp3","Igfbp3","Plac8","Clca1","Klk10","Atp7b","Lgals3","Ptp4a1","Chil4","Krt7","Crip1","Ly6a","AU040972","Lyz2","Klk11","Ptprz1","AA467197")
```
```{r}
P1_sec_automatic_geneList<-DE_P1_secretory_genotype$gene[DE_P1_secretory_genotype$p_val_adj<0.001 & abs(DE_P1_secretory_genotype$avg_logFC)>0.5 & abs(DE_P1_secretory_genotype$pct.1-DE_P1_secretory_genotype$pct.2)>0.15]
```
```{r}
library(ggrepel)
```
```{r}
DE_P1_secretory_genotype$gene<-rownames(DE_P1_secretory_genotype)
#DE_P4_secretory_genotype$sig<-DE_P4_secretory_genotype$p_val_adj<0.001
DE_P1_secretory_genotype$threshold<- ifelse(DE_P1_secretory_genotype$avg_logFC>0 & DE_P1_secretory_genotype$p_val_adj<0.001, "wt_enrich",ifelse(DE_P1_secretory_genotype$avg_logFC<0 & DE_P1_secretory_genotype$p_val_adj<0.001, "mut_enrich","NotSignificant" ) )
volc = ggplot(DE_P1_secretory_genotype, aes(avg_logFC, -log10(p_val_adj))) + #volcanoplot with avg_logFC versus p_val_adj
    geom_point(aes(col=threshold),size=0.2) + #add points colored by significance
  scale_color_manual(values=c("green", "black","magenta"))+
    ggtitle("P1secretory_wt/mut") + geom_text_repel(data=DE_P1_secretory_genotype[DE_P1_secretory_genotype$gene %in% P1_sec_automatic_geneList,], aes(label=gene), point.padding = 1, box.padding = .3) +
  labs(y = expression(-log[10]*" "*"adjusted pvalue"), x = "avg log fold change") + 
  theme(legend.title = element_blank(), legend.position = "top") 
```
```{r,fig.height=8,fig.width=12}
volc
```

```{r}
P1_Dec11_epi@meta.data$specific_type<-mapvalues(P1_Dec11_epi@meta.data$res.0.8,from=c("0","1","2","3","4","5","6","7","8"),to=c("Secretory","CiliaSecretory","Ciliated","Secretory-Krt4","Secretory","Basal","Secretory-Krt4","Basal/Secretory","Ciliated-Foxn4"))
```
```{r,fig.width=5,fig.height=5}
table(P1_Dec11_epi@meta.data$specific_type,P1_Dec11_epi@meta.data$genotype)
```
```{r,fig.width=5,fig.height=5}
ggplot(data=P1_Dec11_epi@meta.data,aes(genotype,fill=specific_type))+ 
    geom_bar(position="fill")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
##### interferons:
```{r,fig.height=16, fig.width=15}
VlnPlot(object = P1_Dec11_epi, features.plot = grep("Ifn",rownames(P1_Dec11_epi@data),value=T), nCol = 3,x.lab.rot = T,point.size.use = 0.2,group.by="cell_type", legend.position = "left")
```
```{r}
df_P1_epi<-FetchData(P1_Dec11_epi,c("Spdef","Creb3l1","Scgb3a2","Scgb1a1","Krt4","Krt13","Foxa3","Aqp3","Aqp4","Aqp5","Gp2","Sostdc1","Smoc2","Krt14","Krt15","Krt5","Rac2","Clic3","res.0.8","res.1.2","genotype","seq_group","specific_type","cell_type","Defb1","Lyz2","Ltf","Sftpa1","Sftpd","Sftpb","Slpi","Lcn2","Pigr","Muc5b","Muc5ac","Chil4","Muc1","Muc2","Muc4","Muc16","Muc20","Lbp","Cd14","Tlr4","Tlr2","Myd88","Ticam1","Itln1","Lgals3","Reg3g","Nod1","Nod2","Ddx58","Ifih1","Dhx58","Ccl5","Cxcl10","Cxcl2","Cxcl1","Pf4","Cxcl12","Cxcl14","Cxcl15","Cxcl16","Cxcl17","Ccl2","Ccl7","Ccl17","Ccl20","Ccl21a","Ccl25","Ccl27a","Ccl28","Cx3cl1","Il10","Tnf","S100a8","S100a9","Il6","Il18","Il1b","Il1rl1","Ccl11","Ccl24","Il33","Il25","Tslp","F2rl1","Retnla","Alox15","Alox5","Gata2","Tgfb2","Tgfb1","Ormdl3","Ptges","Ptgds","Ptgs2","Hpgds","Tbxas1","Areg"))

```
##### MicrobialSensing:
```{r, fig.height=3, fig.width=7}
for (i in c("Lbp","Cd14","Tlr4","Tlr2","Myd88","Ticam1","Itln1","Reg3g","Lgals3","Nod1","Nod2","Ddx58","Ifih1","Dhx58"))
{
pdf(file = paste("Manuscript/MicrobialSensing_genotype/P1/",i,".pdf", sep = ""), width = 6, height = 5)
print(ggplot(df_P1_epi,aes_string(x="genotype",y=i))+facet_grid(.~cell_type)+geom_dotplot(binaxis="y",aes(fill=genotype),binwidth=0.05,stackdir="center",position=position_dodge(0.8), dotsize=0.2)+stat_compare_means(comparisons = list(c("wt", "mut")),method="wilcox.test",size=4,label="p.adj")+ stat_summary(aes(color=genotype),fun.data=mean_sdl, fun.args = list(mult=1), 
                 geom="pointrange",position=position_dodge(0.7))+ theme(axis.text.x = element_text(angle = 45,hjust=1),strip.text.x = element_text(size = 9, colour = "black", angle = 0)))
dev.off()
}
```

##### antimicrobial effectors:
```{r, fig.height=3, fig.width=7}
for (i in c("Muc1","Muc4","Muc16","Muc20","Muc5b","Muc5ac","Muc2","Defb1","Lyz2","Ltf","Sftpa1","Sftpd","Sftpb","Slpi","Lcn2","Pigr","Chil4"))
{
pdf(file = paste("Manuscript/Effectors_genotype/P1/",i,".pdf", sep = ""), width = 6, height = 5)
print(ggplot(df_P1_epi,aes_string(x="genotype",y=i))+facet_grid(.~cell_type)+geom_dotplot(binaxis="y",aes(fill=genotype),binwidth=0.05,stackdir="center",position=position_dodge(0.8), dotsize=0.2)+stat_compare_means(comparisons = list(c("wt", "mut")),method="wilcox.test",size=4,label="p.adj")+ stat_summary(aes(color=genotype),fun.data=mean_sdl, fun.args = list(mult=1), 
                 geom="pointrange",position=position_dodge(0.7))+ theme(axis.text.x = element_text(angle = 45,hjust=1),strip.text.x = element_text(size = 9, colour = "black", angle = 0)))
dev.off()
}
```
##### chemokines:
```{r, fig.height=3, fig.width=7}
for (i in c("Ccl5","Cxcl10","Cxcl2","Cxcl1","Pf4","Cxcl12","Cxcl14","Cxcl15","Cxcl16","Cxcl17","Ccl2","Ccl7","Ccl17","Ccl20","Ccl21a","Ccl25","Ccl27a","Ccl28","Cx3cl1"))
{
pdf(file = paste("Manuscript/chemokines_genotype/P1/",i,".pdf", sep = ""), width = 6, height = 5)
print(ggplot(df_P1_epi,aes_string(x="genotype",y=i))+facet_grid(.~cell_type)+geom_dotplot(binaxis="y",aes(fill=genotype),binwidth=0.05,stackdir="center",position=position_dodge(0.8), dotsize=0.2)+stat_compare_means(comparisons = list(c("wt", "mut")),method="wilcox.test",size=4,label="p.adj")+ stat_summary(aes(color=genotype),fun.data=mean_sdl, fun.args = list(mult=1), 
                 geom="pointrange",position=position_dodge(0.7))+ theme(axis.text.x = element_text(angle = 45,hjust=1),strip.text.x = element_text(size = 9, colour = "black", angle = 0)))
dev.off()
}
```
##### Th2:
```{r, fig.height=3, fig.width=7}
for (i in c("Il10","Tnf","S100a8","S100a9","Il6","Il18","Il1b","Il1rl1","Ccl11","Ccl24","Il33","Il25","Tslp","F2rl1","Retnla","Alox15","Alox5","Gata2","Tgfb2","Tgfb1","Ormdl3","Ptges","Ptgds","Ptgs2","Hpgds","Tbxas1","Areg"))
{
pdf(file = paste("Manuscript/Th2_genotype/P1/",i,".pdf", sep = ""), width = 6, height = 5)
print(ggplot(df_P1_epi,aes_string(x="genotype",y=i))+facet_grid(.~cell_type)+geom_dotplot(binaxis="y",aes(fill=genotype),binwidth=0.05,stackdir="center",position=position_dodge(0.8), dotsize=0.2)+stat_compare_means(comparisons = list(c("wt", "mut")),method="wilcox.test",size=4,label="p.adj")+ stat_summary(aes(color=genotype),fun.data=mean_sdl, fun.args = list(mult=1), 
                 geom="pointrange",position=position_dodge(0.7))+ theme(axis.text.x = element_text(angle = 45,hjust=1),strip.text.x = element_text(size = 9, colour = "black", angle = 0)))
dev.off()
}
```

```{r, fig.height=5, fig.width=10}

ggplot(df_P1_epi,aes(genotype,Muc20))+facet_grid(.~cell_type)+geom_dotplot(binaxis="y",aes(fill=genotype),binwidth=0.05,stackdir="center",position=position_dodge(0.8), dotsize=0.2)+stat_compare_means(comparisons = list(c("wt", "mut")),method="wilcox.test",size=4,label="p.adj")+ stat_summary(aes(color=genotype),fun.data=mean_sdl, fun.args = list(mult=1), 
                 geom="pointrange",position=position_dodge(0.7))+ theme(axis.text.x = element_text(angle = 45,hjust=1),strip.text.x = element_text(size = 9, colour = "black", angle = 0))
```
```{r, fig.height=5, fig.width=10}

ggplot(df_P1_epi,aes(genotype,Cxcl15))+facet_grid(.~cell_type)+geom_dotplot(binaxis="y",aes(fill=genotype),binwidth=0.05,stackdir="center",position=position_dodge(0.8), dotsize=0.2)+stat_compare_means(comparisons = list(c("wt", "mut")),method="wilcox.test",size=4,label="p.adj")+ stat_summary(aes(color=genotype),fun.data=mean_sdl, fun.args = list(mult=1), 
                 geom="pointrange",position=position_dodge(0.7))+ theme(axis.text.x = element_text(angle = 45,hjust=1),strip.text.x = element_text(size = 9, colour = "black", angle = 0))
```
```{r, fig.height=5, fig.width=10}

ggplot(df_P1_epi,aes(genotype,Cxcl17))+facet_grid(.~cell_type)+geom_dotplot(binaxis="y",aes(fill=genotype),binwidth=0.05,stackdir="center",position=position_dodge(0.8), dotsize=0.2)+stat_compare_means(comparisons = list(c("wt", "mut")),method="wilcox.test",size=4,label="p.adj")+ stat_summary(aes(color=genotype),fun.data=mean_sdl, fun.args = list(mult=1), 
                 geom="pointrange",position=position_dodge(0.7))+ theme(axis.text.x = element_text(angle = 45,hjust=1),strip.text.x = element_text(size = 9, colour = "black", angle = 0))
```

```{r, fig.height=5, fig.width=10}
ggplot(df_P1_epi,aes(genotype,Itln1))+facet_grid(.~cell_type)+geom_dotplot(binaxis="y",aes(fill=genotype),binwidth=0.05,stackdir="center",position=position_dodge(0.8), dotsize=0.2)+stat_compare_means(comparisons = list(c("wt", "mut")),method="wilcox.test",size=4,label="p.adj")+ stat_summary(aes(color=genotype),fun.data=mean_sdl, fun.args = list(mult=1), 
                 geom="pointrange",position=position_dodge(0.7))+ theme(axis.text.x = element_text(angle = 45,hjust=1),strip.text.x = element_text(size = 9, colour = "black", angle = 0))
```



```{r}
save(P1_Dec11_epi,file="seu_P1_Dec11mm10.1.2.0_epi.RData")
```


##### For velocyto:
```{r}
P1_Dec11_ID<-read.csv(file="P1_Dec11_ID.csv",header=F,sep=",",stringsAsFactors = F) 

```
```{r}
head(P1_Dec11_ID)

```

```{r}
P1_Dec11_name<-gsub("x","",P1_Dec11_ID)
head(P1_Dec11_name)
```
```{r}
P1_Dec11_name<-gsub(":","_",P1_Dec11_name)
head(P1_Dec11_name)
```
```{r}
TSNE1_Loomorder_P1_epi<-P1_Dec11_epi@dr$tsne@cell.embeddings[match(P1_Dec11_name,rownames(P1_Dec11_epi@dr$tsne@cell.embeddings)),1]
write(TSNE1_Loomorder_P1_epi,"TSNE1_Loomorder_P1_epi.csv",ncolumns=1,sep=",")
```
```{r}
head(TSNE1_Loomorder_P1_epi)
```
```{r}
TSNE2_Loomorder_P1_epi<-P1_Dec11_epi@dr$tsne@cell.embeddings[match(P1_Dec11_name,rownames(P1_Dec11_epi@dr$tsne@cell.embeddings)),2]
write(TSNE2_Loomorder_P1_epi,"TSNE2_Loomorder_P1_epi.csv",ncolumns=1,sep=",")

```
```{r}
res0.8_Loomorder_P1_epi<-P1_Dec11_epi@meta.data$res.0.8[match(P1_Dec11_name,P1_Dec11_epi@cell.names)]
write(res0.8_Loomorder_P1_epi,"res0.8_Loomorder_P1_epi.csv",ncolumns=1,sep=",")

```

































