---
title: "Trachea_WT_10x"
output: html_document
---

```{r}
load("E16_Dec7v3_Trachea.RData") 
```

```{r}
load("seu_P1_Dec11_mm10.1.2.0.RData")
```

```{r}
seu_E16Dec7_P1Dec11<-MergeSeurat(object1 = E16_Dec7v3_Trachea,object2 = seu_P1_Dec11_mm10.1.2.0.RData,min.cells = 1,min.genes = 1,project="E16Dec7_P1Dec11")
```
```{r}
seu_E16Dec7_P1Dec11@meta.data<-seu_E16Dec7_P1Dec11@meta.data[,-which(names(seu_E16Dec7_P1Dec11@meta.data) %in% c("res.0.8", "res.1.2"))]
```
```{r}
load("P4_10X_mm10_1.2.0.RData")
```

```{r}
seu_E16Dec7_P1Dec11_P4Oct18<-MergeSeurat(object1 = seu_E16Dec7_P1Dec11,object2 = P4_10X_mm10_1.2.0,min.cells = 1,min.genes = 1,project="E16Dec7_P1Dec11_P4Oct18")
```
##### remove cluster identities carried over from previous analysis:
```{r}
seu_E16Dec7_P1Dec11_P4Oct18@meta.data<-seu_E16Dec7_P1Dec11_P4Oct18@meta.data[,-which(names(seu_E16Dec7_P1Dec11_P4Oct18@meta.data) %in% c("res.0.8", "res.1.2"))]
```
##### sanity check:
```{r}
table(seu_E16Dec7_P1Dec11_P4Oct18@meta.data$seq_group,seu_E16Dec7_P1Dec11_P4Oct18@meta.data$genotype)
```
```{r}
table(seu_E16Dec7_P1Dec11_P4Oct18@meta.data$seq_group,seu_E16Dec7_P1Dec11_P4Oct18@meta.data$age)
```
```{r}
table(seu_E16Dec7_P1Dec11_P4Oct18@meta.data$seq_group,seu_E16Dec7_P1Dec11_P4Oct18@meta.data$sample_date)
```
```{r}
table(seu_E16Dec7_P1Dec11_P4Oct18@meta.data$seq_group,seu_E16Dec7_P1Dec11_P4Oct18@meta.data$gate)
```
```{r}
table(seu_E16Dec7_P1Dec11_P4Oct18@meta.data$seq_group[is.na(seu_E16Dec7_P1Dec11_P4Oct18@meta.data$doublet_score)])
```

##### get wt cells only:
```{r}
seu_E16Dec7_P1Dec11_P4Oct18 <- SetAllIdent(object = seu_E16Dec7_P1Dec11_P4Oct18, id = "genotype")
seu_E16P1P4_wt<- SubsetData(object=seu_E16Dec7_P1Dec11_P4Oct18,ident.use=c("wt"))
```
```{r}
aggregate(seu_E16P1P4_wt@meta.data[, c(1:2,8:9)], list(seu_E16P1P4_wt@meta.data$seq_group), median)

```
```{r}
seu_E16P1P4_wt <- ScaleData(object = seu_E16P1P4_wt)
```


```{r}
seu_E16P1P4_wt <- FindVariableGenes(object = seu_E16P1P4_wt, do.plot = TRUE, x.low.cutoff=0.1,x.high.cutoff = Inf, y.cutoff = 0.5)
```
```{r}
seu_E16P1P4_wt <- RunPCA(object = seu_E16P1P4_wt,pcs.compute = 40, do.print = FALSE)
seu_E16P1P4_wt <- ProjectPCA(object = seu_E16P1P4_wt, do.print = FALSE)
```
```{r,fig.height=12,fig.width=14}
PCHeatmap(object = seu_E16P1P4_wt, pc.use = c(1:3,18:20,38:40), cells.use = 500, do.balanced = TRUE, label.columns = FALSE, num.genes = 30)

```

```{r}
PCElbowPlot(object = seu_E16P1P4_wt,num.pc = 40)
```
```{r}
n.pcs = 20
res.used <- 0.8

seu_E16P1P4_wt <- FindClusters(object = seu_E16P1P4_wt, reduction.type = "pca", dims.use = 1:n.pcs, 
    resolution = res.used, print.output = 0, save.SNN = TRUE)
```
```{r}
seu_E16P1P4_wt <- RunTSNE(object = seu_E16P1P4_wt, dims.use = 1:n.pcs, seed.use = 10, perplexity=30, dim.embed = 2)

```
```{r}
myRainbow<-c("yellowgreen","yellow","wheat4","wheat1","sandybrown","violetred2","darkgreen","turquoise","turquoise4","red","plum1","purple","indianred1","yellow3","thistle","tan4","olivedrab1","springgreen","snow3","slategray2","slateblue","khaki2","pink3","steelblue1")
```
```{r}
TSNEPlot(object = seu_E16P1P4_wt, do.label = T,pt.size = 0.2)+scale_color_manual(values=myRainbow)

```
```{r}
TSNEPlot(object = seu_E16P1P4_wt, do.label = F,group.by="seq_group",pt.size = 0.2)

```
```{r}
TSNEPlot(object = seu_E16P1P4_wt, do.label = F,group.by="age",pt.size = 0.2)

```


```{r,fig.height=8,fig.width=28}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Trp63","Krt5","Col6a1","Col6a2","Col6a3","Bgn","Postn","Tk1","Mki67","Top2a","Creb3l1","Muc5ac","Gp2","Rac2","Galnt6","Ptgdr","Cd177","Foxj1","Foxn4","Calca","Tubb3","Snap25","Ascl1","Chga","Mbp","Plp1","Mpz","Adipoq","Car3","Egfl7","Cdh5","Pecam1","C1qb","Fcer1g","Cd3g","Acta2","Actc1","Tagln","Alas2","Mia","Wif1","Col8a1","Col11a1","Acan","Col2a1","Sox9","Dcn","Meg3","Clec3b","Adam33","Rarres2","Pi16","Cygb","Pparg","Adrp","Fabp1","Fabp4","Fabp5","Lpl","Lipa","Ednra","Ednrb","Notch3","Ppp1r14a","Myh11","Sfrp2","Mgp","Foxl1","Pdgfra","Twist2","Ano1"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T
  )
```
```{r,fig.height=8,fig.width=28}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Trp63","Krt5","Col6a1","Col6a2","Col6a3","Bgn","Postn","Tk1","Mki67","Top2a","Creb3l1","Muc5ac","Gp2","Rac2","Galnt6","Ptgdr","Cd177","Foxj1","Foxn4","Calca","Tubb3","Snap25","Ascl1","Chga","Mbp","Plp1","Mpz","Adipoq","Car3","Egfl7","Cdh5","Pecam1","C1qb","Fcer1g","Cd3g","Acta2","Actc1","Tagln","Alas2","Mia","Wif1","Col8a1","Col11a1","Acan","Col2a1","Sox9","Dcn","Meg3","Clec3b","Adam33","Rarres2","Pi16","Cygb","Pparg","Adrp","Fabp1","Fabp4","Fabp5","Lpl","Lipa","Ednra","Ednrb","Notch3","Ppp1r14a","Myh11","Sfrp2","Mgp","Foxl1","Pdgfra","Twist2","Ano1"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = F
  )
```



```{r,fig.height=10, fig.width=12}
VlnPlot(object = seu_E16P1P4_wt, features.plot = c("nGene","doublet_score","nUMI","percent.mito","percent.ribo"), nCol = 1,x.lab.rot = T,point.size.use = 0.2)
```
```{r}
GenePlot(object = seu_E16P1P4_wt, gene1 = "Muc5ac", gene2 = "doublet_score", use.raw=F,cex.use = 0.7,cell.ids = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8==12])
```
```{r,fig.height=2,fig.width=4}
GenePlot(object = seu_E16P1P4_wt, gene1 = "Gp2", gene2 = "doublet_score", use.raw=F,cex.use = 0.5,cell.ids = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8==11],col.use = "red")
```
```{r,fig.height=2,fig.width=4}
GenePlot(object = seu_E16P1P4_wt, gene1 = "Gp2", gene2 = "doublet_score", use.raw=F,cex.use = 0.5,cell.ids = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8==12],col.use = "red")
```

```{r}
GenePlot(object = seu_E16P1P4_wt, gene1 = "Muc5ac", gene2 = "doublet_score", use.raw=F,cex.use = 0.7,cell.ids = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8==11])
```



```{r}
library(ggplot2)
```
```{r}
ggplot(data=seu_E16P1P4_wt@meta.data,aes(x=age,group=cell_type,fill=cell_type))+ 
    geom_density(position="fill")+scale_fill_manual(values=myRainbow)
```

```{r}
ggplot(data=seu_E16P1P4_wt@meta.data[seu_E16P1P4_wt@meta.data$age %in% c("E16","P1"),],aes(age,fill=cell_type))+ 
    geom_bar(position="fill")+scale_fill_manual(values=myRainbow)
```
```{r,fig.height=10, fig.width=12}
VlnPlot(object = seu_E16P1P4_wt, features.plot = c("Foxj1","Foxn4"), nCol = 1,x.lab.rot = T,point.size.use = 0.2,group.by = "cell_type")
```

```{r,fig.height=10, fig.width=12}
VlnPlot(object = seu_E16P1P4_wt, features.plot = c("Ano1","Cftr"), nCol = 1,x.lab.rot = T,point.size.use = 0.2,group.by = "cell_type",use.raw=F)
```

```{r}
E16P1P4_wt_res0.8_marker <- FindAllMarkers(object = seu_E16P1P4_wt, only.pos = TRUE, min.pct = 0.25, thresh.use = 0.25)

```

```{r}
E16P1P4_wt_res0.8_marker %>% group_by(cluster) %>% top_n(20, avg_logFC)
```

```{r}
write.table(E16P1P4_wt_res0.8_marker,"marker_res08_E16P1P4wtTrachea.txt",sep="\t")
```



```{r,fig.height=8,fig.width=22}
seu_E16P1P4_wt=buildClusterTree(seu_E16P1P4_wt,do.reorder = F,reorder.numeric = F,pcs.use = 1:20)

```
```{r,fig.height=10,fig.width=18}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Trp63","Krt5","Col6a1","Col6a2","Col6a3","Bgn","Postn","Twist2","Tk1","Mki67","Top2a","Ndufa4l2","Nt5dc2","Igfbp2","Ndrg2","Creb3l1","Muc5ac","Gp2","Rpl9-ps6","Rpl13-ps3","Rps26-ps1","Retnla","Ltf","Ano1","Hmmr"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8 %in% c(23,2,18,20)],group.by="cell_type",cex.row = 20,group.cex = 30
  )
```
```{r}
table(seu_E16P1P4_wt@meta.data$res.0.8,seu_E16P1P4_wt@meta.data$seq_group)
```
```{r}
E16P1P4wt_res08_2over20<-FindMarkers(seu_E16P1P4_wt,ident.1=c(2),ident.2 = c(20),only.pos = TRUE)
E16P1P4wt_res08_2over20
```
```{r}
E16P1P4wt_res08_20over2<-FindMarkers(seu_E16P1P4_wt,ident.1=c(20),ident.2 = c(2),only.pos = TRUE)
E16P1P4wt_res08_20over2
```
```{r}
seu_E16P1P4_wt<-AddMetaData(object = seu_E16P1P4_wt, metadata = seu_E16P1P4_wt@meta.data$res.0.8, col.name = "cell_type")
#seu_E16P1P4_wt@meta.data$cell_type<-seu_E16P1P4_wt@meta.data$res.0.8
```
```{r}
seu_E16P1P4_wt@meta.data$cell_type<-mapvalues(seu_E16P1P4_wt@meta.data$res.0.8,from=c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","23"),to=c("Mesenchyme_E","Mesenchyme_E","Basal_E","Secretory_E","Mesenchyme_P1","Mesenchyme_cycling","Mesenchyme_E","Secretory_P","Mesenchyme_P4","Muscle","Mesenchyme_cycling","Ciliated_P","Ciliated","Mix","Immune","Endothelium","Muscle_P","Mesenchyme_E","Basal_cycling","Mesenchyme","Basal_P","Endothelium","Mesenchyme_P4","Doublet"))
```
```{r,fig.height=5,fig.width=8}
TSNEPlot(object = seu_E16P1P4_wt, do.label = F,group.by="cell_type",pt.size = 0.2)+scale_color_manual(values=myRainbow)

```
```{r,fig.height=8,fig.width=18}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Spdef","Creb3l1","Gp2","Tff2","Scgb1a1","Cyp4b1","Foxj1","Foxn4","Hmmr"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8 %in% c(3,7,11,12)],group.by="cell_type",cex.row = 20,group.cex = 30
  )
```
```{r,fig.height=8,fig.width=10}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Spdef","Creb3l1","Gp2","Tff2","Scgb1a1","Cyp4b1","Foxj1","Foxn4"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8 %in% c(11,12)],group.by="cell_type",cex.row = 30,group.cex = 30
  )
```
```{r,fig.height=8,fig.width=20}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Spdef","Creb3l1","Gp2","Tff2","Scgb1a1","Cyp4b1","Foxj1","Foxn4"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8 %in% c(12)],group.by="age",cex.row = 30,group.cex = 30
  )
```
```{r,fig.height=8,fig.width=20}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Spdef","Creb3l1","Gp2","Tff2","Scgb1a1","Cyp4b1","Foxj1","Foxn4"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8 %in% c(11)],group.by="age",cex.row = 30,group.cex = 30
  )
```

```{r}
E16P1P4wt_res08_11over12<-FindMarkers(seu_E16P1P4_wt,ident.1=c(11),ident.2 = c(12),only.pos = TRUE)
E16P1P4wt_res08_11over12
```

```{r}
cc.genes[! cc.genes %in% mm10.1.2.0_geneList$x]
```
##### need to standardize the gene names in the regev_lab_cell_cycle_genes.txt file, and then load again.
```{r}
cc.genes <- readLines(con = "regev_lab_cell_cycle_genes.txt")
cc.genes<-sub(".*\t","",cc.genes) #### the origninal list concatenate human and mouse names. Need to clean up. I just removed human half of each name
```
```{r}
cc.genes[! cc.genes %in% mm10.1.2.0_geneList$x]
```
##### good all fixed
```{r}
head(cc.genes)
```
```{r}
s.genes <- cc.genes[1:43]
g2m.genes <- cc.genes[44:97]
seu_E16P1P4_wt <- CellCycleScoring(object = seu_E16P1P4_wt, s.genes = s.genes, g2m.genes = g2m.genes, 
    set.ident = TRUE)
```

```{r,fig.height=10, fig.width=12}
VlnPlot(object = seu_E16P1P4_wt, features.plot = c("S.Score","G2M.Score"), nCol = 1,x.lab.rot = T,point.size.use = 0.2,group.by="res.0.8")
```
```{r}
table(seu_E16P1P4_wt@meta.data$Phase,seu_E16P1P4_wt@meta.data$res.0.8)
```
```{r,fig.height=6,fig.width=6}
GenePlot(object = seu_E16P1P4_wt, gene1 = "G2M.Score", gene2 = "S.Score", use.raw=T,cex.use = 0.5,cell.ids = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8==18],group.by="Phase")
```
```{r,fig.height=6,fig.width=6}
GenePlot(object = seu_E16P1P4_wt, gene1 = "G2M.Score", gene2 = "S.Score", use.raw=T,cex.use = 0.5,cell.ids = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8==20],group.by="Phase")
```
```{r,fig.height=6,fig.width=6}
GenePlot(object = seu_E16P1P4_wt, gene1 = "G2M.Score", gene2 = "S.Score", use.raw=T,cex.use = 0.5,cell.ids = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.0.8==2],group.by="Phase")
```

```{r,fig.height=4,fig.width=28}
percentile_table_wt<-apply(seu_E16P1P4_wt@data,1,percent_rank)

```
```{r,fig.height=4,fig.width=28}
percentile_table_wt[1:6,1:6]

```
```{r}
 CellCycle_score_wt<- apply(percentile_table_wt[,colnames(percentile_table_wt) %in% cc.genes],1,mean)
```
```{r}
 head( CellCycle_score_wt)
```
```{r}
seu_E16P1P4_wt<-AddMetaData(object = seu_E16P1P4_wt, metadata = CellCycle_score_wt, col.name = "CellCycle_score")

```
```{r}
VlnPlot(object = seu_E16P1P4_wt, features.plot = c("CellCycle_score"), nCol = 1,x.lab.rot = T,point.size.use = 0.3,use.raw=F,group.by="res.0.8")

```
```{r}
#####Get DE genes between E16 ciliated and goblet to look for difference in Notch signaling (IPA)



```
```{r}
wt_res08_12and3<-FindMarkers(seu_E16P1P4_wt,cells.1<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.0.8==12),cells.2<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.0.8==3),only.pos = F)
wt_res08_12and3
```
```{r}
write.table(wt_res08_12and3,"wt_res08_12and3.txt",sep="\t")
```


#####res.1.2:
```{r}
n.pcs = 20
res.used <- 1.2

seu_E16P1P4_wt <- FindClusters(object = seu_E16P1P4_wt, reduction.type = "pca", dims.use = 1:n.pcs, 
    resolution = res.used, print.output = 0, save.SNN = TRUE,k.param=30)
```
```{r}
seu_E16P1P4_wt <- RunTSNE(object = seu_E16P1P4_wt, dims.use = 1:n.pcs, seed.use = 10, perplexity=30, dim.embed = 2)

```

```{r}
TSNEPlot(object = seu_E16P1P4_wt, do.label = T,pt.size = 0.2)

```
```{r}
TSNEPlot(object = seu_E16P1P4_wt, do.label = T,pt.size = 0.2,group.by="res.1.2")

```



```{r}
TSNEPlot(object = seu_E16P1P4_wt, do.label = F,pt.size = 0.2,group.by="age")

```
```{r}
TSNEPlot(object = seu_E16P1P4_wt, do.label = T,pt.size = 0.2,group.by="cell_type")

```
```{r}
table(seu_E16P1P4_wt@meta.data$res.0.8,seu_E16P1P4_wt@meta.data$res.1.2)
```
```{r,fig.height=8,fig.width=40}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Spdef","Creb3l1","Krt14","Krt5","Trp63","Krt8","Krt18","Gp2","Tff2","Muc5ac","Scgb1a1","Foxj1","Ano1"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,group.by="res.1.2",cex.row = 30,group.cex = 30
  )
```
```{r,fig.height=7,fig.width=30}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Mki67","Top2a","Trp63","Krt5","Krt14","Krt8","Krt18","Spdef","Creb3l1","Gp2","Tff2","Muc5ac","Scgb1a1","Foxj1","Foxn4"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.1.2 %in% c(1,6,3,13,14,21,22,20,25)],group.by="res.1.2",cex.row = 30,group.cex = 30,group.order = c(20,3,13,25,1,6,14,21,22)
  )
```
##### Notch regulated genes in different clusters. c1 is E16 secretory. c6 is P secretory. c3 and c13 E16 basal. c14 E16 ciliated and a few P ciliated. c20 proliferating basal. c21 and 22 P ciliated. 25 P basal. 
```{r,fig.height=8,fig.width=40}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Foxj1","Foxn4","Creb3l1","Spdef","Scgb1a1","Krt5","Epcam","Hes1","Hes6","Myc","Hey1","Hey2","Acta2","Nrarp","Ascl1","Cxcl8","Bcl2","Pten","Notch1","Ccnd1","Heyl"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.1.2 %in% c(1,6,3,13,14,21,22,20,25)],group.order = c(20,3,13,1,14,25,6,21,22),group.by="res.1.2",cex.row = 30,group.cex = 30
  )
```
```{r,fig.height=8,fig.width=40}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Foxj1","Foxn4","Rfx1","Rfx2","Rfx3","Rfx4","Rfx5","Rfx6","Rfx7","Rfx8","Rfx9","Creb3l1","Spdef","Scgb1a1","Krt5","Epcam"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.1.2 %in% c(1,6,3,13,14,21,22,20,25)],group.order = c(20,3,13,1,14,25,6,21,22),group.by="res.1.2",cex.row = 30,group.cex = 30
  )
```
```{r,fig.height=12,fig.width=50}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Foxj1","Foxn4","Cetn2","Plk4","Sas6","Cep152","Ccdc67","Ccdc78","Mcidas","Ccno","Cdc20b","Shisa8","Lrrc23","Sntn","Dnah5","Ccdc39","Rfx2","Ift122","Ift172","Wdpcp","Ttc25","Rfx3","Dync2li1","Dnah9","Dnah11","Creb3l1","Spdef","Scgb1a1","Trp63","Krt5","Epcam","Lrrc10b"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.1.2 %in% c(1,6,3,13,14,21,22,20,25)],group.order = c(20,3,13,1,14,25,6,21,22),group.by="res.1.2",cex.row = 30,group.cex = 30
  )
```
```{r,fig.height=12,fig.width=50}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Foxj1","Foxn4","Cetn2","Plk4","Sas6","Cep152","Ccdc67","Ccdc78","Mcidas","Ccno","Cdc20b","Shisa8","Lrrc23","Sntn","Dnah5","Ccdc39","Rfx2","Ift122","Ift172","Wdpcp","Ttc25","Rfx3","Dync2li1","Dnah9","Dnah11","Creb3l1","Spdef","Scgb1a1","Trp63","Krt5","Epcam","Lrrc10b"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = F,cells.use=seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.1.2 %in% c(6,21,22)],group.order = c(6,21,22),group.by="res.1.2",cex.row = 30,group.cex = 30
  )
```


```{r}
wt_res08_21and22<-FindMarkers(seu_E16P1P4_wt,cells.1<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.2==21),cells.2<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.2==22),only.pos = F)
wt_res08_21and22
```
```{r}
wt_res08_22over6and21<-FindMarkers(seu_E16P1P4_wt,cells.1<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.2==22),cells.2<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.2 %in% c(6,21)),only.pos = T)
wt_res08_22over6and21
```
```{r}
wt_res08_22over21<-FindMarkers(seu_E16P1P4_wt,cells.1<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.2==22),cells.2<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.2 %in% c(21)),only.pos = T)

```
```{r}
wt_res08_22over6<-FindMarkers(seu_E16P1P4_wt,cells.1<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.2==22),cells.2<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.2 %in% c(6)),only.pos = T)

```
```{r}
intersect(rownames(wt_res08_22over21),rownames(wt_res08_22over6))
```
```{r,fig.height=12,fig.width=50}

DoHeatmap(object = seu_E16P1P4_wt, genes.use =intersect(rownames(wt_res08_22over21),rownames(wt_res08_22over6)), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.1.2 %in% c(1,6,3,13,14,21,22,20,25)],group.order = c(20,3,13,1,14,25,6,21,22),group.by="res.1.2",cex.row = 30,group.cex = 30
  )
```
```{r}
VlnPlot(object = seu_E16P1P4_wt, features.plot = c("doublet_score"), nCol = 1,x.lab.rot = T,point.size.use = 0.3,use.raw=F,group.by="res.1.2")

```


```{r,fig.height=8,fig.width=20}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Spdef","Creb3l1","Krt14","Krt5","Trp63","Krt8","Krt18","Ano1"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use = seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.1.2 %in% c(3,13)],group.by="res.1.2",cex.row = 30,group.cex = 30
  )
```
##### res.1.4:
```{r}
n.pcs = 20
res.used <- 1.4

seu_E16P1P4_wt <- FindClusters(object = seu_E16P1P4_wt, reduction.type = "pca", dims.use = 1:n.pcs, 
    resolution = res.used, print.output = 0, save.SNN = TRUE,k.param = 15)
```
```{r}
seu_E16P1P4_wt <- RunTSNE(object = seu_E16P1P4_wt, dims.use = 1:n.pcs, seed.use = 10, perplexity=30, dim.embed = 2)

```

```{r}
TSNEPlot(object = seu_E16P1P4_wt, do.label = T,pt.size = 0.2,group.by="res.1.4")

```
```{r,fig.width=10,fig.height=5}
TSNEPlot(object = seu_E16P1P4_wt, do.label = F,pt.size = 0.2,group.by="cell_type")+scale_color_manual(values=myRainbow)

```
```{r}
TSNEPlot(object = seu_E16P1P4_wt, do.label = F,pt.size = 0.2,group.by="seq_group",cells.use=seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$age=="P4"],colors.use = c("green","magenta"))

```
```{r}
TSNEPlot(object = seu_E16P1P4_wt, do.label = F,group.by="seq_group",pt.size = 0.2,,cells.use=seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$seq_group %in% c("P4_Oct18_wt_green","P4_Oct18_wt_red","P1_Dec11_wt_6")],colors.use = c("blue","green","magenta"))

```
```{r, fig.height=5, fig.width=12}
FeaturePlot(seu_E16P1P4_wt, c("nGene","nUMI","Epcam"), pt.size = 0.3, nCol = 3,no.legend = F)
```



```{r,fig.height=8,fig.width=28}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Trp63","Krt5","Col6a1","Col6a2","Col6a3","Bgn","Postn","Tk1","Mki67","Top2a","Creb3l1","Muc5ac","Gp2","Rac2","Galnt6","Ptgdr","Cd177","Foxj1","Foxn4","Calca","Tubb3","Snap25","Ascl1","Chga","Mbp","Plp1","Mpz","Adipoq","Car3","Egfl7","Cdh5","Pecam1","C1qb","Fcer1g","Cd3g","Acta2","Actc1","Tagln","Alas2","Mia","Wif1","Col8a1","Col11a1","Acan","Col2a1","Sox9","Dcn","Meg3","Clec3b","Adam33","Rarres2","Pi16","Cygb","Pparg","Adrp","Fabp1","Fabp4","Fabp5","Lpl","Lipa","Ednra","Ednrb","Notch3","Ppp1r14a","Myh11","Sfrp2","Mgp","Foxl1","Pdgfra","Twist2","Ano1"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,group.by="res.1.4"
  )
```

```{r,fig.height=7,fig.width=30}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Mki67","Top2a","Trp63","Krt5","Krt14","Krt8","Krt18","Spdef","Creb3l1","Krt4","Krt13","Gp2","Tff2","Muc5ac","Scgb1a1","Foxj1","Foxn4"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.1.4 %in% c(1,14,33,3,12,31,16,20,23,28,4,35)],group.by="res.1.4",cex.row = 30,group.cex = 30,group.order = c(33,1,14,3,12,31,16,28,35,4,23,20)
  )
```

```{r}
wt_res14_3over12<-FindMarkers(seu_E16P1P4_wt,cells.1<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.4==3),cells.2<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.4==12),only.pos = T)
wt_res14_3over12
```

```{r,fig.height=8,fig.width=22}
seu_E16P1P4_wt<-SetAllIdent(seu_E16P1P4_wt,id="res.1.4")
seu_E16P1P4_wt=buildClusterTree(seu_E16P1P4_wt,do.reorder = F,reorder.numeric = F,pcs.use = 1:20)

```

```{r,fig.height=15,fig.width=30}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Col6a1","Bgn","Postn","Tk1","Mki67","Top2a","Calca","Tubb3","Snap25","Ascl1","Chga","Mbp","Plp1","Mpz","Adipoq","Car3","Egfl7","Cdh5","Pecam1","Nrp1","Lyve1","C1qb","Fcer1g","Cd3g","Acta2","Actc1","Tagln","Alas2","Mia","Wif1","Col8a1","Col11a1","Acan","Col2a1","Sox9","Dcn","Meg3","Col14a1","Clec3b","Adam33","Rarres2","Pi16","Cygb","Plin2","Fabp1","Fabp4","Fabp5","Lpl","Lipa","Ednra","Ednrb","Notch3","Ppp1r14a","Myh11","Sfrp2","Mgp","Foxl1","Pdgfra","Twist2","Ano1","Prdm16","Ucp1","Pparg","Ng2","Pdgfrb","Cd34","Ly6a","Zfp423","Lepr"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[!(seu_E16P1P4_wt@meta.data$res.1.4 %in% c(1,14,33,3,12,31,16,20,23,28,4,35,32))],group.by="res.1.4",cex.row = 20,group.cex = 30,group.order=c(21,30,18,15,19,7,5,9,10,22,25,17,26,27,34,24,29,8,0,2,11,6,13)
  )
```
```{r,fig.height=15,fig.width=30}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Egfl7","Cdh5","Pecam1","Nrp1","Lyve1","Acta2","Tagln","Myh11","Ppp1r14a","Ednrb","Notch3","Mki67","Top2a","C1qb","Fcer1g","Cd3g","Snap25","Ascl1","Chga","Calca","Mbp","Plp1","Mpz","Adipoq","Car3","Pi16","Lpl","Mia","Col11a1","Acan","Col2a1","Wif1","Mgp","Postn","Twist2"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[!(seu_E16P1P4_wt@meta.data$res.1.4 %in% c(1,14,33,3,12,31,16,20,23,28,4,35,32))],group.by="res.1.4",cex.row = 25,group.cex = 30,group.order=c(21,30,18,10,9,5,15,19,7,22,25,27,17,26,34,24,29,8,2,6,11,0,13)
  )
```

```{r,fig.height=20,fig.width=45}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Egfl7","Cdh5","Pecam1","Nrp1","Lyve1","Mia","Col11a1","Acan","Col2a1","Wif1","Mgp","Postn","Acta2","Tagln","Myh11","Ppp1r14a","Ednrb","Notch3","Pi16","Lpl","Mki67","Top2a","Snap25","Ascl1","Chga","Calca","Mbp","Plp1","Mpz","C1qb","Fcer1g","Cd3g","Adipoq","Car3","Twist2"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[!(seu_E16P1P4_wt@meta.data$res.1.4 %in% c(1,14,33,3,12,31,16,20,23,28,4,35,32))],group.by="cell_type",cex.row = 30,group.cex = 30,group.order=c("Endothelial_blood-vessel","Endothelial_lymphatic-vessel","pChondrocyte","pFibroblast_1","pMuscle","pFibroblast_2","pAdipocyte","pFibroblast_3","pCyclingFibroblast","Neural","Immune","Mix","eChondrocyte","eFibroblast_1","eMuscle","eFibroblast_2","eFibroblast_3","eCyclingFibroblast")
  )
```


##### Remove cells of high doublet scores:
```{r,fig.height=20,fig.width=45}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Egfl7","Cdh5","Pecam1","Nrp1","Lyve1","Mia","Col11a1","Acan","Col2a1","Wif1","Mgp","Postn","Acta2","Tagln","Myh11","Ppp1r14a","Ednrb","Notch3","Pi16","Lpl","Mki67","Top2a","Snap25","Ascl1","Chga","Calca","Mbp","Plp1","Mpz","C1qb","Fcer1g","Cd3g","Adipoq","Car3","Twist2"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[!(seu_E16P1P4_wt@meta.data$res.1.4 %in% c(1,14,33,3,12,31,16,20,23,28,4,35,32)) & seu_E16P1P4_wt@meta.data$doublet_score<0.2],group.by="cell_type",cex.row = 30,group.cex = 30,group.order=c("Endothelial_blood-vessel","Endothelial_lymphatic-vessel","pChondrocyte","pFibroblast_1","pMuscle","pFibroblast_2","pAdipocyte","pFibroblast_3","pCyclingFibroblast","Neural","Immune","Mix","eChondrocyte","eFibroblast_1","eMuscle","eFibroblast_2","eFibroblast_3","eCyclingFibroblast")
  )
```

```{r,fig.height=15,fig.width=30}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Mia","Wif1","Col8a1","Col11a1","Acan","Col2a1","Comp","Matn1","Mmp13","Col10a1","Sox9","Dcn","Meg3","Col14a1","Clec3b","Adam33","Rarres2","Pi16","Cygb","Adrp","Sox9","Creb3l2","Nkx3-1","Nkx3-2","Rarg","Runx2"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[!(seu_E16P1P4_wt@meta.data$res.1.4 %in% c(1,14,33,3,12,31,16,20,23,28,4,35))],group.by="res.1.4",cex.row = 20,group.cex = 30,group.order=c(18,15,21,30,17,26,27,34,24,19,29,8,22,25,10,5,7,9,0,2,11,6,13,32)
  )
```
```{r,fig.height=15,fig.width=30}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Adipoq","Car3","Fabp4","Fabp5","Lpl","Lipa","Myh11","Acta2","Actc1","Tagln","Ano1","Prdm16","Ucp1","Pparg","Ng2","Pdgfrb","Cd34","Ly6a","Zfp423","Lepr","Dlk1"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[!(seu_E16P1P4_wt@meta.data$res.1.4 %in% c(1,14,33,3,12,31,16,20,23,28,4,35))],group.by="res.1.4",cex.row = 20,group.cex = 30,group.order=c(18,15,21,30,17,26,27,34,5,10,7,11,24,19,29,8,0,9,2,22,25,6,13,32)
  )
```



```{r}
wt_res14_10over5<-FindMarkers(seu_E16P1P4_wt,cells.1<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.4==10),cells.2<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.4==5),only.pos = T)
wt_res14_10over5
```

```{r}
wt_res14_21over30<-FindMarkers(seu_E16P1P4_wt,cells.1<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.4==21),cells.2<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.4==30),only.pos = T)
wt_res14_21over30
```
```{r}
wt_res14_30over21<-FindMarkers(seu_E16P1P4_wt,cells.1<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.4==30),cells.2<-WhichCells(object=seu_E16P1P4_wt,cells.use = seu_E16P1P4_wt@meta.data$res.1.4==21),only.pos = T)
wt_res14_30over21
```
```{r,fig.height=15,fig.width=30}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Egfl7","Cdh5","Pecam1","Nrp1","Lyve1","Acta2","Tagln","Myh11","Ppp1r14a","Ednrb","Notch3","Mki67","Top2a","C1qb","Fcer1g","Cd3g","Snap25","Ascl1","Chga","Calca","Mbp","Plp1","Mpz","Adipoq","Car3","Pi16","Lpl","Mia","Col11a1","Acan","Col2a1","Wif1","Mgp","Postn","Twist2"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[(seu_E16P1P4_wt@meta.data$res.1.4 %in% c(27))],group.by="res.1.4",cex.row = 20,group.cex = 30
  )
```
```{r,fig.height=15,fig.width=30}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Adipoq","Car3","Pi16","Cygb","Pparg","Adrp","Fabp1","Fabp4","Fabp5","Lpl","Lipa","Cd3g","Snap25","Tubb3","Chga","Plp1","Mbp"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[(seu_E16P1P4_wt@meta.data$res.1.4 %in% c(26))],group.by="seq_group",cex.row = 20,group.cex = 30
  )
```
```{r,fig.height=15,fig.width=30}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Adipoq","Car3","Pi16","Cygb","Pparg","Adrp","Fabp1","Fabp4","Fabp5","Lpl","Lipa","Cd3g","Snap25","Tubb3","Chga","Plp1","Mbp"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[(seu_E16P1P4_wt@meta.data$res.1.4 %in% c(27))],group.by="seq_group",cex.row = 20,group.cex = 30
  )
```

```{r}
seu_E16P1P4_wt@meta.data$cell_type<-mapvalues(seu_E16P1P4_wt@meta.data$res.1.4,from=c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","23","24","25","26","27","28","29","30","31","32","33","34","35"),to=c("eFibroblast_1","eBasal-Krt14","eChondrocyte","eSecretory-Krt4","pSecretory","eFibroblast_2","eFibroblast_1","pFibroblast_2","eFibroblast_3","eFibroblast_2","eFibroblast_2","pChondrocyte","eSecretory","pCyclingFibroblast","eBasal","pMuscle","eCiliated","Immune","eMuscle","pFibroblast_2","pMucusCiliated","Endothelial_blood-vessel","eCyclingFibroblast","pCiliated","pFibroblast_1","eCyclingFibroblast","Neural","Mix","pBasal","pFibroblast_3","Endothelial_lymphatic-vessel","eCyclingSecretory","Doublet","eCyclingBasal","pAdipocyte","pCyclingSecretory"))
```

```{r,fig.height=10,fig.width=34}

DoHeatmap(object = seu_E16P1P4_wt, genes.use = c("Epcam","Mki67","Top2a","Trp63","Krt5","Krt14","Krt8","Krt18","Spdef","Creb3l1","Krt4","Gp2","Muc5ac","Scgb1a1","Foxj1","Foxn4"), 
    slim.col.label = TRUE, group.label.rot = TRUE,use.scaled = T,cells.use=seu_E16P1P4_wt@cell.names[seu_E16P1P4_wt@meta.data$res.1.4 %in% c(1,14,33,3,12,31,16,20,23,28,4,35)],group.by="cell_type",cex.row = 30,group.cex = 30,group.order=c("eCyclingBasal","eBasal-Krt14","eBasal","eSecretory-Krt4","eSecretory","eCyclingSecretory","eCiliated","pBasal","pSecretory","pCyclingSecretory","pCiliated","pMucusCiliated")
  )
```
```{r,fig.width=10,fig.height=4}
VlnPlot(object = seu_E16P1P4_wt, features.plot = c("doublet_score"), nCol = 1,x.lab.rot = T,point.size.use = 0.3,use.raw=F,group.by="cell_type")

```
```{r}
VlnPlot(object = seu_E16P1P4_wt, features.plot = c("CellCycle_score"), nCol = 1,x.lab.rot = T,point.size.use = 0.3,use.raw=F,group.by="cell_type")

```
```{r}
wt_mocosaGoblet_score<- apply(percentile_table_wt[,colnames(percentile_table_wt) %in% OMIMgene_mucosaGoblet],1,mean)
```
```{r}
 max( wt_mocosaGoblet_score)
```
```{r}
seu_E16P1P4_wt<-AddMetaData(object = seu_E16P1P4_wt, metadata = wt_mocosaGoblet_score, col.name = "mocosaGoblet_score")

```
```{r}
VlnPlot(object = seu_E16P1P4_wt, features.plot = c("mocosaGoblet_score"), nCol = 1,x.lab.rot = T,point.size.use = 0.3,use.raw=F,group.by="cell_type")

```
```{r}
 wt_ciliopathy_table<- percentile_table_wt[,colnames(percentile_table_wt) %in% OMIMgene$Ciliopathy]
```

```{r}
 wt_ciliopathy_score<- apply(wt_ciliopathy_table,1,mean)
```
```{r}
 head(wt_ciliopathy_score)
```
```{r}
 wt_PCD_score<- apply(percentile_table_wt[,colnames(percentile_table_wt) %in% OMIMgene$Primary.ciliary.dyskinesia],1,mean)
```
```{r}
 head(wt_PCD_score)
```
```{r}
seu_E16P1P4_wt<-AddMetaData(object = seu_E16P1P4_wt, metadata = wt_ciliopathy_score, col.name = "ciliopathy_score")

```
```{r}
seu_E16P1P4_wt<-AddMetaData(object = seu_E16P1P4_wt, metadata = wt_PCD_score, col.name = "PCD_score")

```
```{r}
VlnPlot(object = seu_E16P1P4_wt, features.plot = c("ciliopathy_score"), nCol = 1,x.lab.rot = T,point.size.use = 0.3,use.raw=F,group.by="cell_type")

```

```{r}
VlnPlot(object = seu_E16P1P4_wt, features.plot = c("PCD_score"), nCol = 1,x.lab.rot = T,point.size.use = 0.3,use.raw=F,group.by="cell_type")

```
```{r}
save(seu_E16P1P4_wt,file="seu_E16P1P4_wt.RData")
```





















































































































































